Plots for Latent Hawkes Paper

### 2020-11-15 output_scripts/Paper_Plots.Rmd
adding plots for fig 4 repeated with data simulated from m1 and m2

## Figure 4a Recover Latent Ranks
```{r fig-4a}
sim_data_path <- "../output/sims_long/"

@@

```{r prepare data for 4b,cache=TRUE}
plot.s <- 9#7#9
plot.i <- 4
plot.j <- 5

@@

```{r fig4b-recover-latent-states}
y.ub <- c(20,20,20)#c(130,72,31)
x_events <- 135#150
#model_colors <- c("steelblue", "goldenrod2", "firebrick2")
model_colors <- viridis::viridis(3)
layout(matrix(c(1:6), 6, 1), heights=rep(c(2.5,1),3))
layout(matrix(c(1:6), 6, 1), heights=rep(c(7.5,1),3))
my.xlim <- 140#160#200#130
my.long.x <- 140#205#134.5
legend.x <- c(50,50,50)#c(135,135,135)
legend.y <- #c(34,32,32)
legend.y <- c(20,12,15)#c(62,37,18)#c(22,22,22)
legend.cex <- 1.5 #3
line.alpha <- c(0.9,0.6)
line.wdth <- c(2,2.5)
negative.col <- add.alpha('lightskyblue',0.6)
positive.col <- add.alpha('tomato',0.6)
#negative.col <- viridis::viridis(5,alpha = 0.6)[1]
#positive.col <- viridis::viridis(5,alpha = 0.6)[4]
par(mar = c(0,0.3,0,0.1), oma = c(3,3.5,4,0),
    tcl=0.2,mgp=c(0.5,0,0), xpd=TRUE)

@@

legend(10,legend.y[2],c("State 0/1 events","State change point"),
       col = c(NA,"red"),
       y.intersp=0.85,x.intersp=-0.1, bty = "n",
       pch = c(NA,4), pt.cex = c(NA,2), cex=legend.cex,
       lty = c(NA,NA), lwd=c(NA,4))
points(14,8,pch=1,cex=2,col="blue")
points(16,8,pch=16,cex=2,col="blue")

@@

>>>>

## Figure 4a Recover Latent Ranks
```{r fig-4a}
sim_data_path <- "../output/sims/sims_m3_dc/"

@@

```{r prepare data for 4b,cache=TRUE}
plot.s <- 1#7#9
plot.i <- 4
plot.j <- 5

@@

```{r fig4b-recover-latent-states}
y.ub <- c(17,17,17)#c(130,72,31)
x_events <- 135#150
#model_colors <- c("steelblue", "goldenrod2", "firebrick2")
model_colors <- viridis::viridis(3)
layout(matrix(c(1:6), 6, 1), heights=rep(c(2.5,1),3))
layout(matrix(c(1:6), 6, 1), heights=rep(c(7.5,1),3))
my.xlim <- 140#160#200#130
my.long.x <- 140#205#134.5
legend.x <- c(50,50,50)#c(135,135,135)
legend.y <- #c(34,32,32)
legend.y <- c(15,12,15)#c(62,37,18)#c(22,22,22)
legend.cex <- 1.5 #3
line.alpha <- c(0.9,0.6)
line.wdth <- c(2,2.5)
negative.col <- add.alpha('lightskyblue',0.6)
positive.col <- add.alpha('tomato',0.6)
#negative.col <- viridis::viridis(5,alpha = 0.6)[1]
#positive.col <- viridis::viridis(5,alpha = 0.6)[4]
par(mar = c(0,0.3,0,0.1), oma = c(3,3.5,4,0),
    tcl=0.2,mgp=c(0.5,0,0), xpd=TRUE)

@@

legend(10,legend.y[2]+2,c("State 0/1 events","State change point"),
       col = c(NA,"red"),
       y.intersp=0.85,x.intersp=-0.1, bty = "n",
       pch = c(NA,4), pt.cex = c(NA,2), cex=legend.cex,
       lty = c(NA,NA), lwd=c(NA,4))
points(14,8+2,pch=1,cex=2,col="blue")
points(16,8+2,pch=16,cex=2,col="blue")

@@

## Fig 4 repeated for data generated from other models

For completeness, we also wish to repeat the above two plots using data generated 
from M1 and M2.

For data simulated from M1.

```{r est ranks for m1 sim data}
m1_path <- "../output/sims/M1/"

m1_files <- list.files(m1_path, pattern = "sim_model1_fit*")
#m1_files

# load(paste0(m1_path,m1_files[1]))
sim_model1_stan_sim1_total <- list()
sim_model1_stan_sim2_total <- list()
sim_model1_stan_sim3_total <- list()

for(i in seq_along(m1_files)) {
  load(paste0(m1_path,m1_files[i]))
  sim_model1_stan_sim1_total[[i]] <- sim_model1_stan_sim1
  sim_model1_stan_sim2_total[[i]] <- sim_model1_stan_sim2
  sim_model1_stan_sim3_total[[i]] <- sim_model1_stan_sim3

}

m2_path <- "../output/sims/M2/"

m2_files <- list.files(m2_path, pattern = "sim_model2_fit*")
#m1_files

# load(paste0(m1_path,m1_files[1]))
sim_model2_stan_sim1_total <- list()
sim_model2_stan_sim2_total <- list()
sim_model2_stan_sim3_total <- list()

for(i in seq_along(m2_files)) {
  load(paste0(m2_path,m2_files[i]))
  sim_model2_stan_sim1_total[[i]] <- sim_model2_stan_sim1
  sim_model2_stan_sim2_total[[i]] <- sim_model2_stan_sim2
  sim_model2_stan_sim3_total[[i]] <- sim_model2_stan_sim3

}

```{r plot est ranks M1 and M2}
n_sim <- 50

    ## Posterior density
    current_result <- eval(parse(text=paste("sim_model1_stan_sim",
                                            m,
                                            "_total",
                                            sep="")))
    for(i in c(1:n_sim)){
      lines(density(current_result[[i]]$f[1001:2000,j]),
            col=add.alpha(model_colors[m], alpha=0.8),cex=0.6,lwd=0.8)
    }

    if(m==3){
      text(text.x.adj[j],text.y[j],text.x[j],cex=2.4)
    }
  }
  ## bottom
  mtext(text=expression(f[1],f[2],f[3],f[4],f[5]), side=1, line=2.4, outer=TRUE, 
        at=c(1:length(object_par$f_vec_1))/length(object_par$f_vec_1)-0.1,
        cex=2.4)
  mtext(text="Latent rank variables", side=1, line=5.3, outer=TRUE, cex=2.7,
        at=0.5)
  ## top
  mtext(text="Simulated from M1", side=3, line=0, outer=TRUE, cex=3,
        at=0.5, font=2)
  ## left
  mtext(text=rev(c("C-HP  ","C-DCHP","C-MMHP")), side=2, line=0.5, outer=TRUE, 
        at=c(0:2)/3+0.15,
        cex=2, las=2)
}

    ## Posterior density
    current_result <- eval(parse(text=paste("sim_model2_stan_sim",
                                            m,
                                            "_total",
                                            sep="")))
    for(i in c(1:n_sim)){
      lines(density(current_result[[i]]$f[1001:2000,j]),
            col=add.alpha(model_colors[m], alpha=0.8),cex=0.6,lwd=0.8)
    }

    if(m==3){
      text(text.x.adj[j],text.y[j],text.x[j],cex=2.4)
    }
  }
  ## bottom
  mtext(text=expression(f[1],f[2],f[3],f[4],f[5]), side=1, line=2.4, outer=TRUE, 
        at=c(1:length(object_par$f_vec_1))/length(object_par$f_vec_1)-0.1,
        cex=2.4)
  mtext(text="Latent rank variables", side=1, line=5.3, outer=TRUE, cex=2.7,
        at=0.5)
  ## top
  mtext(text="Simulated from M2", side=3, line=0, outer=TRUE, cex=3,
        at=0.5, font=2)
  ## left
  mtext(text=rev(c("C-HP  ","C-DCHP","C-MMHP")), side=2, line=0.5, outer=TRUE, 
        at=c(0:2)/3+0.15,
        cex=2, las=2)
}
```

Similarly, we can then repeat this for Fig 4b also for both simulated 
data examples, with slightly more difficulty.

```{r process m1 for fig4b}
plot.s <- 1#7#9
plot.i <- 4
plot.j <- 5

cur_i <- plot.i
cur_j <- plot.j

load(paste0(m1_path,m1_files[plot.s]))

temp.t <- sim_model1_data[[1]]$day_hour[ which(sim_model1_data[[1]]$start == cur_i & 
                  sim_model1_data[[1]]$end == cur_j)]
# temp.t <- test.mmhp$tau
current.n <- length(temp.t)-1
time.segment <- seq(0,tail(temp.t,1),length.out=5000)

### Preprocess the model inference result
## model 1
#########
model1_par_est <- list(lambda0 =
                         mean(sim_model1_stan_sim1_total[[plot.s]]$lambda0[1001:2000]),
                       eta_1=mean(sim_model1_stan_sim1_total[[plot.s]]$eta_1[1001:2000]),
                       eta_2=mean(sim_model1_stan_sim1_total[[plot.s]]$eta_2[1001:2000]),
                       eta_3=mean(sim_model1_stan_sim1_total[[plot.s]]$eta_3[1001:2000]),
                       beta=mean(sim_model1_stan_sim1_total[[plot.s]]$beta[1001:2000]),
                       f=apply(sim_model1_stan_sim1_total[[plot.s]]$f[1001:2000,],2,mean))
lambda.m1<-uniHawkesIntensityNumeric(object=list(lambda0 = 
                                                   model1_par_est$lambda0,
                                    alpha =
                                      model1_fn$alpha.fun(model1_par_est$f[plot.i],
                                    model1_par_est$f[plot.j],
                                    model1_par_est$eta_1,
                                    model1_par_est$eta_2,
                                    model1_par_est$eta_3),
                                    beta=model1_par_est$beta),
                                     events=temp.t,
                                     time.vec=time.segment)

## model2
#############
model2_par_est <- list(gamma=apply(sim_model1_stan_sim2_total[[plot.s]]$gamma[1001:2000,],
                                   2,mean),
                    zeta = 
                      apply(sim_model1_stan_sim2_total[[plot.s]]$zeta[1001:2000,],2,mean),
                       eta_1=mean(sim_model1_stan_sim2_total[[plot.s]]$eta_1[1001:2000]),
                       eta_2=mean(sim_model1_stan_sim2_total[[plot.s]]$eta_2[1001:2000]),
                       eta_3=mean(sim_model1_stan_sim2_total[[plot.s]]$eta_3[1001:2000]),
                       beta=mean(sim_model1_stan_sim2_total[[plot.s]]$beta[1001:2000]),
                       f=apply(sim_model1_stan_sim2_total[[plot.s]]$f[1001:2000,],2,mean))

## Model3
#############
mmhp_par_est <- list(lambda0=mean(sim_model1_stan_sim3_total[[plot.s]]$lambda0[1001:2000]),
                     lambda1=mean(sim_model1_stan_sim3_total[[plot.s]]$lambda1[1001:2000]),
                     eta_1=mean(sim_model1_stan_sim3_total[[plot.s]]$eta_1[1001:2000]),
                     eta_2=mean(sim_model1_stan_sim3_total[[plot.s]]$eta_2[1001:2000]),
                     eta_3=mean(sim_model1_stan_sim3_total[[plot.s]]$eta_3[1001:2000]),
                     beta=mean(sim_model1_stan_sim3_total[[plot.s]]$beta[1001:2000]),
                     f=apply(sim_model1_stan_sim3_total[[plot.s]]$f[1001:2000,],2,mean))

### need to load in the interpolated latent state here
#### TO DO ####

lambda.true <- mmhpTrueIntensityNumeric(params=object_true,
                                        t=temp.t,
                       latent=list(x=fixStateTransition(test.mmhp)$x,
                              z=fixStateTransition(test.mmhp)$z),
                              time.vec=time.segment)
```
